---
title: "Assignment 5 - Data Report: U.S. Shootings in 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Information

```{r, include = FALSE}
source("analysis.R")
```
  This report analyzes a dataset containing information about _U.S. shootings in 2018_ ([source](http://www.shootingtracker.com/)). According to the dataset, there were **`r number_of_shootings` shootings**, with  **`r number_lost` people killed in the United States**. When analyzing the data by location, **the U.S. city that was most impacted was `r most_impacted_city`**, with impact being quantified by the number of casualties. **In `r most_impacted_city`, `r most_impact_city_numbers` people were either killed or injured.** When analyzing the data by U.S. states, **`r num_states_with_shootings` states had shootings in 2018**, and **`r get_correct_grammar(state_most_shootings)` had the most shootings**. **The state with the most casualties was `r state_most_casualties` with `r number_casualties` casualties.**
  
## Summary Table

```{r, echo = FALSE}
kable(agg_table,
      col.names = c("State", "Number of Shootings",
                    "Injured", "Killed", "Total Casualties", "% of Casualties Killed")) %>%
  kable_styling(bootstrap_options = "striped", fixed_thead = T) %>% 
  scroll_box(width = "100%", height = "400px")
```

In the table above, the data is sorted by U.S. State and contains the _number of shootings in each state_, _number of people injured_, _number of people killed_, _the total number of casualties in 2018_, and _the percentage of people who died when compared to the total number of casualties_. The table is arranged by total number of shootings, so the state with the most shootings appears at the top. It should be noted that **not all states are listed within the table, but only the `r num_states_with_shootings` states with recorded shootings in 2018**. The table depicts `r get_correct_grammar(state_most_shootings)` to have the most shootings; additionally, _states with a higher number of shootings have more overall casualties_. `r state_most_casualties` has the highest recorded number of casualties and is also the state with the most shootings. On the other hand, _when looking that the percentage of casualties that were killed, the data stays more consistent across all states_, with the average being `r avg_percent`; however, there are outliers to this average. An extreme value within the dataset is the percent of casualties killed in `r max_state`, in which `r max_percent` of the casualties had died.
