---
title: "Assignment 5 - Data Report: U.S. Shootings in 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Information

```{r, include = FALSE}
source("analysis.R")
```
  This report analyzes a dataset containing information about _U.S. shootings in 2018_ ([source](http://www.shootingtracker.com/)). According to the dataset, there were **`r number_of_shootings` shootings**, with  **`r number_lost` people killed in the United States**. When analyzing the data by location, **the U.S. city that was most impacted was `r most_impacted_city`**, with impact being quantified by the number of casualties. **In `r most_impacted_city`, `r most_impact_city_numbers` people were either killed or injured.** When analyzing the data by U.S. states, **`r num_states_with_shootings` states had shootings in 2018**, and **`r get_correct_grammar(state_most_shootings)` had the most shootings**. **The state with the most casualties was `r state_most_casualties` with `r number_casualties` casualties.**
  
## Summary Table

```{r, echo = FALSE}
kable(agg_table,
      col.names = c("State", "Number of Shootings",
                    "Injured", "Killed", "Total Casualties", "% of Casualties Killed")) %>%
  kable_styling(bootstrap_options = "striped", fixed_thead = T) %>%
  scroll_box(width = "100%", height = "400px")
```

<br>
In the table above, the data is sorted by U.S. State and contains the _number of shootings in each state_, _number of people injured_, _number of people killed_, _the total number of casualties in 2018_, and _the percentage of people who died when compared to the total number of casualties_. The table is arranged by total number of shootings, so the state with the most shootings appears at the top. It should be noted that **not all states are listed within the table, but only the `r num_states_with_shootings` states with recorded shootings in 2018**. The table depicts `r get_correct_grammar(state_most_shootings)` to have the most shootings; additionally, _states with a higher number of shootings have more overall casualties_. `r state_most_casualties` has the highest recorded number of casualties and is also the state with the most shootings. On the other hand, _when looking that the percentage of casualties that were killed, the data stays more consistent across all states_, with the average being `r avg_percent`; however, there are outliers to this average. An extreme value within the dataset is the percent of casualties killed in `r max_state`, in which `r max_percent` of the casualties had died.

## Particular Incident: Delaware - July 9th, 2018
A particular incident of this dataset is a shooting that occured on _`r date_case`_ in _`r case_city`, `r case_state`_, resulting in `r num_killed_case` people killed and `r num_injured_case` survivors. This particular case involved a murder-suicide, where a husband of a family had shot his family and then himself. Their neighbor had claimed that the couple had outspoken marital issues ([source](https://www.nbcphiladelphia.com/news/national-international/shooting-wilmington-delaware-home/2065392/)). This incident is the only recorded shooting in Delaware in 2018.

## Interactive Visualization - Shootings in the U.S. in 2018
_Purpose: The interactive map below visualizes the distribution of shootings by location, with the circles weighted by the number of people killed in the shooting._
`r interactive_map`

_Looking at the map above,_ we see that shootings are **more concentrated in the East than in the Western United States**. In particular, there is a **dense number of shootings in the Northeast**. This doesn't particular mean that there are a high number of shootings in a particular state in this region--on the contrary, the states with the highest number of shootings are outside this region (`r get_correct_grammar(state_most_shootings)`). Additionally, **there are very sparse, if any, shootings in the middle region of the U.S.**, and **there are more concentrated shootings along the East and West coasts**. 

## Plot
This plot intends to bring insight on _what months of the year had the most shootings in 2018_.

```{r, echo = FALSE}
months_barplot
```

<br> 
As seen _in the bar plot above_, **the most shootings occurred in `r month_most_shootings` (`r month_most_number`), while the least number of shootings occurred in `r month_least_shootings` (`r month_least_number`).** When regarding the overall trend of the graph, **there is an increase in shootings in the warmer months, while in cooler months there is less**. The increase seen in the warmer months is dramatic as well, with the **percent increase of shootings from `r month_least_shootings` to `r month_most_shootings` being `r perc_increase`%**.
<br><br>